Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-02-06 Report/ieee_journal.tex
halfcooked
adhithan
half cooked
introduction,  convergence distributed
first final report
updated-final darft without optimal word
final_modified wsrm
nothing great

\documentclass[10pt,final,twocolumn,a4paper]{./../Styles/IEEEtran}

\usepackage{amsmath,amssymb}
\usepackage{acronym,etoolbox}
\usepackage[ruled]{algorithm2e}
\usepackage[dvips]{graphicx}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[subrefformat=parens,labelformat=parens,labelseparator=period,caption=false]{subfig}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

@@

%\section{Response to reviewer comments}
%\input{review_comments}

@@

\input{decomposition_solutions}

\subsection{Decomposition based Approaches} \label{sec-4a}

\subsubsection{Primal Decomposition Approach} \label{sec-4.1}

\input{primal_approach}
%\input{primal_algorithm}

\subsubsection{\ac{ADMM} approach} \label{sec-4.2}

\input{dual_approach}
%\input{dual_algorithm}

\subsection{Decomposition via \ac{KKT} Conditions for \ac{MSE} Formulation} \label{sec-4.3}

\input{kkt_mse_approach}

@@

\subsection{Average Backlogged Packets Over Slots}

@@

\subsection{Convergence Analysis} \label{sec-3.5}
\input{convergence_centralized}

>>>>

\documentclass[10pt,final,twocolumn,letterpaper]{./../Styles/IEEEtran}

\usepackage{amsmath,amssymb}
\usepackage{acronym,etoolbox}
\usepackage[ruled]{algorithm2e}
\usepackage[dvipdfmx]{graphicx}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[subrefformat=parens,labelformat=parens,labelseparator=period,caption=false]{subfig}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

@@

\section{Response to reviewer comments}
\input{review_comments}

@@

\review{\input{decomposition_solutions}}

%\subsection{Decomposition based Approaches} \label{sec-4a}

\subsection{Primal Decomposition} \label{sec-4.1}
\input{primal_approach}
%\input{primal_algorithm}

\subsection{Alternating Directions Method of Multipliers (ADMM)} \label{sec-4.2}
\input{dual_approach}
%\input{dual_algorithm}

\subsection{Convergence Analysis for Distributed Algorithms} \label{sec-dist-conv}
\review{\input{convergence_distributed}}

\subsection{Decomposition via \ac{KKT} Conditions for \ac{MSE} Formulation} \label{sec-4.3}
\input{kkt_mse_approach}

@@

\subsection{Average Backlogged Packets Over Slots} \label{time-correlated}

@@

\section{Convergence Analysis of Centralized Algorithm} \label{sec-3.5}
\review{\input{convergence_centralized}}

### 2015-02-06 Report/review_comments.tex
nothing great

\resp We thank the reviewer for reading the paper and providing the comments. Please consider the responses in line with the comments.

@@

\resp We fully agree with the reviewer comment. The Q-WSRM scheme is throughput optimal when the queues associated with the users are significantly large in comparison with the transmission rate (service rate). In order to restrict the over allocation of the available resources to a specific user beyond the total number of queued packets, we have included the additional rate constraint in the Q-WSRME algorithm. It would be ideal to compare with the Q-WSRME algorithm which performs similar to the Q-WSRM when the queued packets are large enough to be emptied by the current transmissions. Fig. \ref{fig-review} compares the centralized algorithms using average number of backlogged packets after \me{100} slots of transmission for different arrival rate. Note that the average arrival rate of all the users are same but the instantaneous arrivals are based on the Poisson distribution. It can be seen from Fig. \ref{fig-review} that the JSFRA scheme with \me{q=2} performs similar to the Q-WSRME scheme for higher arrival rates, which explains the dropping of the squared rate term in the Q-WSRM formulation. When the average arrival rate is smaller, the JSFRA scheme with \me{q=2} performs noticeably better than the Q-WSRME scheme as can be seen from the average number of backlogged packets. In all packet arrival scenarios, JSFRA scheme with \me{q=1} reduces the average number of backlogged packets over the given slots is lesser than the Q-WSRME scheme.
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{Review/reviewer.eps}
\label{fig-review}
\caption{System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace}}
\end{figure}

@@

\resp We understand the reviewer concerns. We have modified the convergence section to include additional information to provide more clarity. \review{Update converge proof}

@@

\resp We thank the reviewer for the raising the concern on the selection order. We have elaborated the discussion on the sub-channel wise selection scheme. \review{discuss on selection}

@@

\resp We thank the reviewer for the insightful comment on the practicality of the distributed algorithm. We fully agree with the reviewer comment on the information exchange between the BSs in the distributed approach. Please note that the proposed distributed algorithm based on the primal or dual decomposition is provided for the completeness purpose. Note that the distributed approach is performed for the convex subproblem, which leads to the same stationary point asymptotically as that of the centralized solution. In reality, we have to limit the number of iterations required for each distributed algorithm, thereby leading to a point which is not the stationary point when the algorithm is allowed to converge. In section \ref{sec-4.3}, we have discussed a practical approach based on the KKT conditions, which attains a sub-optimal point with few number of iterations.

@@

\resp We understand the reviewers concern. We have update the text on the convergence of the distributed algorithm (ADMM). Please not that the ADMM or the primal decomposition algorithm is used for the convex subproblem only. If the distributed algorithm is iterated until convergence, it is guaranteed to attain the same stationary point as that of the centralized algorithm \cite{boyd2011distributed}.

>>>>

\resp We thank the reviewer for reading the paper and providing the comments. The response for the reviewer questions are made in line.

@@

\resp We fully agree with the reviewer comment. The Q-WSRM scheme is throughput optimal when the queues associated with the users are significantly large in comparison with the transmission rate (service rate). In order to restrict the over allocation of the available resources to a specific user beyond the total number of queued packets, we have included the additional rate constraint in the Q-WSRME algorithm. It would be ideal to compare with the Q-WSRME algorithm which performs similar to the Q-WSRM when the queued packets are large enough to be emptied by the current transmissions. Fig. \ref{fig-review} compares the centralized algorithms using average number of backlogged packets after \me{100} slots of transmission for different arrival rate. Note that the average arrival rate of all the users are same but the instantaneous arrivals are based on the Poisson distribution. It can be seen from Fig. \ref{fig-review} that the JSFRA scheme with \me{q=2} performs similar to the Q-WSRME scheme for higher arrival rates, which explains the dropping of the squared rate term in the Q-WSRM formulation. When the average arrival rate is smaller, the JSFRA scheme with \me{q=2} performs noticeably better than the Q-WSRME scheme as can be seen from the average number of backlogged packets. In all packet arrival scenarios, JSFRA scheme with \me{q=1} reduces the average number of backlogged packets over the given slots is lesser than the Q-WSRME scheme.
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{Review/reviewer.eps}
\label{fig-review}
\caption{System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 4,2,12,1 \rbrace}}
\end{figure}
The discussions are included in the Section \ref{time-correlated}.

@@

\resp We understand the reviewer concerns. We have modified the convergence section to include additional information to provide more clarity. Appendix \ref{sec-3.5} discuss the convergence proof of the proposed algorithm in detail.

@@

\resp We thank the reviewer for the raising the concern on the selection order. We have elaborated the discussion on the sub-channel wise selection scheme in Section \ref{sec-3.4}.

@@

\resp We thank the reviewer for the insightful comment on the practicality of the distributed algorithm. We fully agree with the reviewer comment on the information exchange between the BSs in the distributed approach. Please note that the proposed distributed algorithm based on the primal or dual decomposition is provided for the completeness purpose. Note that the distributed approach is performed for the convex subproblem, which leads to the same stationary point asymptotically as that of the centralized solution. In reality, we have to limit the number of iterations required for each distributed algorithm, thereby leading to a point which is not the stationary point when the algorithm is allowed to converge. In Section \ref{sec-4.3}, we have discussed a practical approach based on the KKT conditions, which attains a sub-optimal point with few number of iterations.

@@

\resp We understand the reviewers concern. We have update the text on the convergence of the distributed algorithm (ADMM). Please not that the ADMM or the primal decomposition algorithm is used for the convex subproblem only. If the distributed algorithm is iterated until convergence, it is guaranteed to attain the same stationary point as that of the centralized algorithm \cite{boyd2011distributed}. We have included this discussion in Section \ref{sec-dist-conv}.

### 2015-02-06 Report/simulation_results.tex
Updated WSRM
nothing great

We discuss the performance of the centralized algorithms in Section \ref{sec-3} for some system configurations. To begin with, we consider a single cell \ac{SISO} model operating at \me{10} dB \ac{SNR} with \me{K = 3} users sharing \me{N = 3} sub-channel resources. The number of packets waiting at the transmitter for each user is given by \me{Q_k = 4,8} and \me{4} bits, respectively. 
\begin{table*}
	\centering
		\caption{Sub-channel-wise listing of channel gains and rate allocations by different algorithms for a scheduling instant}
	\renewcommand{\arraystretch}{1.25} \scriptsize
	\begin{tabular}{|*{14}{c|}}
		\hline
		\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRME approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
		\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(modified \review{\emph{backpressure}})} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
		\cline{3-14}
		&& SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} \\
		\hline
		\me{1} & \me{4} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
		\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.88}  &  \me{3.11} & \me{0} &  \me{5.49}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
		\me{3} & \me{4} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{4.0} & \me{5.81} &  \me{0}  &  \me{0} \\
		\hline
		\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{3.92} bits} & \multicolumn{3}{c|}{\me{2.51} bits} & \multicolumn{3}{c|}{\me{5.89} bits} \\
		\hline
	\end{tabular}
	\label{tbl-1} \vspace{-0.1in}
\end{table*}
\begin{figure*}
	\centering
	\subfloat[][System Model \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 4,3,9,4,1\rbrace}]{
		\includegraphics[width=0.5\textwidth]{fig-1-4}
		\label{fig-1}}
	\subfloat[][System Model \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 2,3,9,4,2\rbrace}]{
		\includegraphics[width=0.5\textwidth]{fig-2-5}
		\label{fig-2}}
	\caption{Total number of backlogged packets \me{\chi} present in the system after each \ac{SCA} updates}
	\label{fig-a} \vspace{-0.1in}
\end{figure*}

Table \ref{tbl-1} tabulates the channel seen by the users over each sub-channel followed by the rates assigned by three different algorithms, \ac{Q-WSRME} allocation, \ac{JSFRA} approach and the band-wise \ac{Q-WSRM} scheme using the \ac{WMMSE} design \cite{wmmse_shi}. The performance metric used for the comparison is the total number of backlogged bits left over at each slot after the allocation, which is denoted as \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. Even though \me{\mc{U}(1)} and \me{\mc{U}(3)} has equal number of backlogged packets of \me{Q_{1} = Q_{3} = 4} bits, user \me{\mc{U}(3)} is scheduled in the first sub-channel due to the better channel condition. In contrast, the \ac{JSFRA} approach assigns the first user on the first sub-channel, which reduces the total number of backlogged packets waiting at the transmitter. The rate allocated for \me{\mc{U}(2)} on the second sub-channel is higher in \ac{JSFRA} scheme compared to the other schemes. It is due to the efficient allocation of the total power shared across the sub-channels.
\begin{table}
	\centering
	\caption{Number of backlogged bits associated with each user for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}.}
	\renewcommand{\arraystretch}{1.25} \scriptsize
	\begin{tabular}{|c|*{8}{c}|c|}
		\hline
		\multirow{2}{*}{\me{q}} & \multicolumn{8}{c|}{user indices} & \multirow{2}{*}{\me{\chi}} \\
		\cline{2-9}
		& 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \\
		\hline
		\hline
		\me{1} & 15.0 & 3.95 & 5.26 & 8.95 & 7.0 & 11.9 & 12.0 & 9.7 & 25.15 \\
		\me{2} & 11.2 & 3.9 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.9 & 27.77 \\
		\me{\infty} & 11.4 & 4.4 & 10.4 & 10.4 & 10.4 & 8.4 &  8.4 &  6.4 & 28.68 \\
		\hline
		\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\
		\cline{1-9}
	\end{tabular}
	\label{tbl-3} \vspace{-0.1in}
\end{table}

@@

For both scenarios in Fig. \ref{fig-a}, the \ac{Q-WSRME} performs marginally inferior to the \ac{JSFRA} algorithms due to the weights used in the algorithm. The performance loss is  attributed to the fact that the \ac{Q-WSRME} algorithm favors the users with the large number of backlogged packets as compared to the users with better channel conditions. Fig. \subref*{fig-2} compares the algorithms for \me{N_R = 2} receive antenna case. In all figures, the receivers are updated along with the \ac{SCA} update instants \textit{i.e}, \me{J_{\max} = 1} in Algorithm \ref{algo-1}. It is also noted that the performance degradation by performing the group update is very minimal. Since the receiver minimizes the objective for the fixed transmit precoders, the convergence is monotonic as can be seen from the figures.

>>>>

We discuss the performance of the centralized algorithms in Section \ref{sec-3} for some system configurations. To begin with, we consider a single cell \ac{SISO} model operating at \me{10} dB \ac{SNR} with \me{K = 3} users sharing \me{N = 3} sub-channel resources. The number of packets waiting at the transmitter for each user is given by \me{Q_k = 4,8} and \me{4} bits, respectively. 
\begin{table*}
	\centering
		\caption{Sub-channel-wise listing of channel gains and rate allocations by different algorithms for a scheduling instant}
	\renewcommand{\arraystretch}{1.25} \scriptsize
	\begin{tabular}{|*{14}{c|}}
		\hline
		\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRME approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
		\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(modified \review{\emph{backpressure}})} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
		\cline{3-14}
		&& SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} \\
		\hline
		\me{1} & \me{4} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
		\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.88}  &  \me{3.11} & \me{0} &  \me{5.49}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
		\me{3} & \me{4} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{4.0} & \me{5.81} &  \me{0}  &  \me{0} \\
		\hline
		\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{3.92} bits} & \multicolumn{3}{c|}{\me{2.51} bits} & \multicolumn{3}{c|}{\me{5.89} bits} \\
		\hline
	\end{tabular}
	\label{tbl-1}
\end{table*}
\begin{figure*}
	\centering
	\subfloat[][System Model \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 4,3,9,4,1\rbrace}]{
		\includegraphics[width=0.5\textwidth]{fig-1-4}
		\label{fig-1}}
	\subfloat[][System Model \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 2,3,9,4,2\rbrace}]{
		\includegraphics[width=0.5\textwidth]{fig-2-5}
		\label{fig-2}}
	\caption{Total number of backlogged packets \me{\chi} present in the system after each \ac{SCA} updates}
	\label{fig-a}
\end{figure*}

Table \ref{tbl-1} tabulates the channel seen by the users over each sub-channel followed by the rates assigned by three different algorithms, \ac{Q-WSRME} allocation, \ac{JSFRA} approach and the band-wise \ac{Q-WSRM} scheme using the \ac{WMMSE} design \cite{wmmse_shi}. The performance metric used for the comparison is the total number of backlogged bits left over at each slot after the allocation, which is denoted as \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. Even though \me{\mc{U}(1)} and \me{\mc{U}(3)} has equal number of backlogged packets of \me{Q_{1} = Q_{3} = 4} bits, user \me{\mc{U}(3)} is scheduled in the first sub-channel due to the better channel condition. In contrast, the \ac{JSFRA} approach assigns the first user on the first sub-channel, which reduces the total number of backlogged packets waiting at the transmitter. The rate allocated for \me{\mc{U}(2)} on the second sub-channel is higher in \ac{JSFRA} scheme compared to the other schemes. It is due to the efficient allocation of the total power shared across the sub-channels.
\begin{table}
	\centering
	\caption{Number of backlogged bits associated with each user for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}.}
	\renewcommand{\arraystretch}{1.25} \scriptsize
	\begin{tabular}{|c|*{8}{c}|c|}
		\hline
		\multirow{2}{*}{\me{q}} & \multicolumn{8}{c|}{user indices} & \multirow{2}{*}{\me{\chi}} \\
		\cline{2-9}
		& 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \\
		\hline
		\hline
		\me{1} & 15.0 & 3.95 & 5.26 & 8.95 & 7.0 & 11.9 & 12.0 & 9.7 & 25.15 \\
		\me{2} & 11.2 & 3.9 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.9 & 27.77 \\
		\me{\infty} & 11.4 & 4.4 & 10.4 & 10.4 & 10.4 & 8.4 &  8.4 &  6.4 & 28.68 \\
		\hline
		\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\
		\cline{1-9}
	\end{tabular}
	\label{tbl-3}
\end{table}

@@

For both scenarios in Fig. \ref{fig-a}, the \ac{Q-WSRME} performs marginally inferior to the \ac{JSFRA} algorithms due to the weights used in the algorithm. The performance loss is  attributed to the fact that the \ac{Q-WSRME} algorithm favors the users with the large number of backlogged packets as compared to the users with better channel conditions. Fig. \subref*{fig-2} compares the algorithms for \me{N_R = 2} receive antenna case. In all figures, the receivers are updated along with the \ac{SCA} update instants \textit{i.e}, \me{J_{\max} = 1} in Algorithm \ref{algo-1}. It is also noted that the degradation by performing combined update is marginal, since the receiver minimizes the objective for a fixed transmit precoders thereby leading to a monotonic convergence.
